---
title: "Detailed results for instances"
description: |
  This report contains links til detailed results for each instance.
author:
  - name: Nicolas Forget
    url: http://pure.au.dk/portal/en/nforget@econ.au.dk
    affiliation: CORAL, BSS, Aarhus University
    affiliation_url: https://econ.au.dk/coral
  - name: Lars Relund Nielsen
    url: http://pure.au.dk/portal/en/larsrn@econ.au.dk
    affiliation: CORAL, BSS, Aarhus University
    affiliation_url: https://econ.au.dk/coral
  - name: Sune Lauth Gadegaard
    url: http://pure.au.dk/portal/en/sgadegaard@econ.au.dk
    affiliation: CORAL, BSS, Aarhus University
    affiliation_url: https://econ.au.dk/coral
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
citation_url: https://mcdmsociety.github.io/MOrepo-Forget20/report_details.html
---

<style type="text/css">
table td, table th {  /* Table  */
  font-size: 10px !important;
}
</style>

```{r setup, include=FALSE}
#' Function for loading missing packages that install them if not already installed.
#'
#' @param packages String vector with package names
#'
#' @return NULL (invisible)
#' @export
#'
#' @examples loadPackages(c("MASS", "ggplot2", "tikzDevice"))
loadPackages <- function(packages) {
  newP <- packages[!(packages %in% installed.packages()[,"Package"])]
  if(length(newP)) install.packages(newP, repos = "http://cran.rstudio.com/")
  lapply(packages, library, character.only = TRUE)
  invisible(NULL)
}
loadPackages(c("tidyverse", "knitr", "rgl", "gMOIP", "rmarkdown", "ggplot2", "plotly", "DT", "RColorBrewer", "wesanderson", "kableExtra", "ggthemes"))

if (isTRUE(getOption('knitr.in.progress'))) options(rgl.useNULL=TRUE)
rgl::setupKnitr()
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, message=FALSE, include = TRUE, 
  cache = TRUE, autodep = TRUE,
  echo=FALSE,
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.7,
  layout="l-page"   #"l-screen-inset"
)
knit_hooks$set(webgl = hook_webgl, rgl = hook_rgl)
options(knitr.kable.NA = '')
```

Detailed results for each instance can be generated using `instance.Rmd` in the subfolder `results/report`. The report with detailed results and plots is already generated for some of the instances: 

```{r}
datAll <- read_csv("../statistics.csv")
instanceFiles <- list.files("../../docs/instances", ".html", full.names = F)
instances <- str_remove(instanceFiles, ".html")

findInstances <- function(dat) {
  instances <- NULL
  dat <- dat %>% 
    filter(solved == 1) %>% 
    group_by(instance) %>%
    summarise_at(vars(contains(c("YN", "total"))), list(mean = mean, sd = sd, max = max), na.rm = TRUE)
  instances <- bind_rows(instances, 
                         tibble(class = "lowND", 
                                instance = dat %>% arrange(YN_max) %>% slice(1:2) %>% pull(instance)))
  instances <- bind_rows(instances, 
                       tibble(class = "highND", 
                              instance = dat %>% arrange(desc(YN_max)) %>% slice(1:2) %>% pull(instance)))
  instances <- bind_rows(instances, 
                         tibble(class = "highNDus", 
                                instance = dat %>% arrange(desc(YN_max - YNs_max)) %>% slice(1:2) %>% pull(instance)))
  instances <- bind_rows(instances, 
                         tibble(class = "highNDse", 
                                instance = dat %>% arrange(desc(YNse_max)) %>% slice(1:2) %>% pull(instance)))
  instances <- bind_rows(instances, 
                         tibble(class = "highNDusPct", 
                                instance = dat %>% arrange(desc((YN_max - YNs_max)/YN_max)) %>% 
                                  slice(1:2) %>% pull(instance)))
  instances <- bind_rows(instances, 
                         tibble(class = "highNDsne", 
                                instance = dat %>% arrange(desc(YNs_max - YNse_max)) %>% slice(1:2) %>% pull(instance)))
  instances <- bind_rows(instances, 
                         tibble(class = "highSDcpu", 
                                instance = dat %>% arrange(desc(tpstotal_sd)) %>% slice(1:2) %>% pull(instance)))
}
datSpecial <- bind_rows(
  findInstances(datAll),
  findInstances(datAll %>% filter(pb == "AP")),
  findInstances(datAll %>% filter(pb == "UFLP")),
  findInstances(datAll %>% filter(pb == "KP")),
  findInstances(datAll %>% 
    filter(coef == "spheredown", rangeC == "[1,1000]" | rangeC == "[1,1000]|[1,100]", varsel == "mof"))
) %>% 
  distinct() %>% 
  arrange(class) 


for (ins in datSpecial %>% filter(!(instance %in% instances)) %>% pull(instance)) {
   reset <- FALSE
   cat("File", ins, "\n")
   try(rmarkdown::render("instance.Rmd", output_file = paste0(ins, ".html"),
                     output_dir = "../../docs/instances", quiet = T, envir = new.env(),
          params=list(new_title=paste("Results for instance", ins) , currentInstance = ins) ))
}
```


```{r, echo=FALSE}
toLink <- function(inst) {
  if (length(inst) == 0) return("")
  links <- str_c('../../docs/instances/', inst, '.html')
  url <- str_c('instances/', inst, '.html')
  if_else(file.exists(links), str_c('<a href="', url, '">', inst, '</a>'), inst)
}


datInput <- datAll %>% 
  distinct(instance, .keep_all = TRUE) %>% 
  select(instance, pb, n, p, coef, ratioNDcoef) %>% 
  filter(instance %in% instances) %>% 
  mutate(instance = toLink(instance))

DT::datatable(
  datInput,
  escape = F,
  options = list(scrollX = T),
  caption = "Instances with detailed results and plots generated. A link is given in the instance column."
)
```

Some instances that might be of interest:

  * Instances with a low number of non-dominated points: `r toLink(datSpecial %>% filter(class = "lowND") %>% pull(instance))`
  * Instances with a high number of non-dominated points: `r toLink(datSpecial %>% filter(class = "highND") %>% pull(instance))`
  * Instances with a high number of unsupported non-dominated points: `r toLink(datSpecial %>% filter(class = "highNDus") %>% pull(instance))`
  * Instances with a high number of supported extreme non-dominated points: `r toLink(datSpecial %>% filter(class = "highNDse") %>% pull(instance))`
  * Instances with a high unsupported non-dominated point percentage: `r toLink(datSpecial %>% filter(class = "highNDusPct") %>% pull(instance))`
  * Instances with a high number of supported non-extreme non-dominated points: `r toLink(datSpecial %>% filter(class = "highNDsne") %>% pull(instance))`
  * Instances with highest variance in cpu time: `r toLink(datSpecial %>% filter(class = "highSDcpu") %>% pull(instance))`














